---
title: "Graphic for communication"
author: "Pikachu"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggrepel)
```

# 28.1 Introduction

The purpose of a plot title is to summarize the main finding. Avoid titles that just describe what the plot is: "A sccater plot of engine displacement and vs fuel economy"

## 28.2.1 Exercises
>1. Create one plot on the fuel economy data with customised title, subtitle, caption, x, y, and colour labels.

```{r}
mpg %>% 
  ggplot(aes(displ, hwy, color = class)) + 
  geom_point() +
  labs(x = 'Displacement (L)',
       y = 'Highway MPG',
       color = 'car type',
       title = 'Fuel economy generally deacreases with eninge size',
       subtitle = '2 seaters car (sport cars) are exceptional because of their lightweight',
       caption = 'Data from fueleconomy.gov')
```

>2. The `geom_smooth()` is somewhat misleading because the `hwy` for large engines is skewed upwards due to the inclusion of lightweight sports cars with big engines. Use your modelling tools to fit and display a better model.

```{r}
mpg %>% 
  ggplot(aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(method = 'lm') +
  labs(x = 'displacement (L)',
       y = 'highway MPG',
       title = 'Fuel economy generally decreases with engine size',
       caption = 'data from fueleconomy.gov') + 
  theme_bw()
```

# 28.3 Annotations
## 28.3.1 Exercises
>1. Use `geom_text()` with infinite positions to place text at the four corners of the plot.

```{r}
label <- tribble(
  ~displ, ~hwy, ~label, ~vjust, ~hjust,
  Inf, Inf, "Top right", "top", "right",
  Inf, -Inf, "Bottom right", "bottom", "right",
  -Inf, Inf, "Top left", "top", "left",
  -Inf, -Inf, "Bottom left", "bottom", "left"
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label, vjust = vjust, hjust = hjust), data = label)
```

>2. How do labels with `geom_text()` interact with faceting? How can you add a label to a single facet? How can you put a different label in each facet? (Hint: think about the underlying data.)

If the facet label is not specified, the text is drawn in all facets

```{r}
mpg %>% 
  ggplot(aes(displ, hwy)) + 
  geom_point() + 
  annotate(label = 'Ignore this', x = Inf, y =Inf, geom = 'text', hjust = 1, vjust = 1) + 
  facet_wrap(~ class)
```
To draw the label in only one facet, add a column to the label data frame with the value of the faceting variable(s) in which to draw it.

```{r}
label <- tibble(class = 'compact', label = 'compact car')

mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_point() + 
  geom_text(data = label, aes(label = label), x = Inf, y = Inf, hjust = 1, vjust = 1) + 
  facet_wrap(~ class)
```
To draw labels in different plots, simply have the facetting variable(s):

```{r}
label <- mpg %>% distinct(class) %>% mutate(label = str_c('label for ', class))

mpg %>% 
  ggplot(aes(displ, hwy)) + 
  geom_point() +
  geom_text(data = label, aes(label = label), x = Inf, y = Inf, hjust = 1, vjust = 1) +
  facet_wrap(~ class)
```


# 28.4 Scales

>3. Change the display of the presidential terms by:
1. Combining the two variants shown above.
2. Improving the display of the y axis.
3. Labelling each term with the name of the president.
4. Adding informative plot labels.
5. Placing breaks every 4 years (this is trickier than it seems!).

```{r}
presidential %>% ggplot(aes(
  name %>% fct_reorder(start),
  start,
  xend = name,
  yend = end,
  color = party
)) + geom_point() + 
  geom_segment() + 
  scale_color_manual('Party', values = c(Democratic = 'blue', Republican = 'red')) +
  labs(x = NULL,
       y = 'year',
       title = 'Terms of US presidents',
       subtitle = 'Roosevelth (34th) to Obama (44th)') +
  scale_y_date(date_breaks = '4 years', date_labels = '%y') +
  coord_flip()
```

# 28.5 Zoomings

# 28.6 Themes

# 28.7 Saving your plots

# 28.8 Learning more