---
title: "Data transformation"
author: "Pikachu"
date: "11/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
```

# 10.1 Introduction
# 10.2 Filter

It’s common to only want to explore one part of a dataset. A great data
analysis strategy is to start with just one observation unit (one person, one
city, etc), and understand how it works before attempting to generalise the
conclusion to others. This is a great technique if you ever feel overwhelmed
by an analysis: zoom down to a small subset, master it, and then zoom back
out, to apply your conclusions to the full dataset.

Filtering is also useful for extracting outliers. Generally, you don’t want
to just throw outliers away, as they’re often highly revealing, but it’s useful
to think about partitioning the data into the common and the unusual. You
summarise the common to look at the broad trends; you examine the outliers
individually to see if you can figure out what’s going on.

For example, look at this plot that shows how the x and y dimensions of
the diamonds are related:

```{r}
diamonds %>% 
  ggplot(aes(x, y)) +
  geom_bin2d() +
  theme_bw()
```
There are around 50,000 points in this dataset: most of them lie along the
diagonal, but there are a handful of outliers. One clear set of incorrect values
are those diamonds with zero dimensions. We can use `filter()` to pull them
out

In a real analysis, you’d look at the outliers in more detail to see if you
can find the root cause of the data quality problem. In this case, we’re just
going to throw them out and focus on what remains.

```{r}
# Remove diamonds with x = 0 or y = 0 or y > 20
diamonds %>% 
  filter(x > 0, y > 0, y < 20) %>%
  ggplot(aes(x, y)) +
  geom_bin2d() + 
  # add regression line
  geom_smooth(method = 'lm', se = F, color = 'red', size = 2, alpha = .5) + 
  theme_bw()
```
This plot is informative. We can see a strong relationship between `x` and `y`. The regression line tells us that for most diamonds, `x` and `y` are very similar. However, this plot sitll has problems:
- The plot is mostly empty, because most of the data lies along the diagonal
- There are some clear bivariate outliers, but it’s hard to select them with a simple filter.

## 10.2.1 Useful tools
See `?filter` 
## 10.2.2 Missing values
`filter()` only includes observations where all arguments are ``TRUE`, so `NA`
values are automatically dropped (because almost any operations with `NA` will return `NA`. If you want to include missing values, be explicit: `x > 10 | is.na(x)`.

## 10.2.3 Exercises
1. Practice your filtering skills by:
• Finding all the diamonds with equal x and y dimensions.
```{r}
diamonds %>% filter(x == y)
```

• A depth between 55 and 70.
```{r}
diamonds %>% filter(depth %>% between(55, 70))
# equivalent
#dimands %>% filter(55 <= depth, depth <= 70)
```

• A carat smaller than the median carat.
```{r}
diamonds %>% filter(carat < mean(carat))
```

• Cost more than $10,000 per carat
```{r}
diamonds %>% filter(price > 10000)
```

• Are of good or better quality
```{r}
diamonds %>% filter(cut %in% c('Good', 'Very Good'))
```

3. Repeat the analysis of outlying values with `z`. Compared to `x` and `y`, how would you characterise the relationship of `x` and `z`, or `y` and `z`?

- `x` and `z`
```{r}
diamonds %>% 
  ggplot(aes(x, z)) +
  geom_bin2d() +
  theme_bw()
```
```{r}
# Remove outliers: x = 0 or y = 0 or z > 10
diamonds %>% 
  filter(x > 0, y > 0, z < 10) %>%
  ggplot(aes(x, z)) +
  geom_bin2d() + 
  geom_smooth(method = 'lm', size = 2, color = 'red') + 
  theme_bw()
```
**Comment**: There is a strong positive relationship between `x` and `z`. But the relationship is weaker than between `x` and `y`, and the variance is larger

- `y` and `z`

```{r}
diamonds %>% 
  ggplot(aes(y, z)) + 
  geom_bin2d() + 
  theme_bw()
```
```{r}
# Remove outliers: y > 20 or z > 10 or y = 0 or z = 0
diamonds %>% 
  filter(y < 20, z < 10, 0 < z, 0 < y) %>%
  ggplot(aes(y, z)) +
  geom_bin2d() +
  geom_smooth(method = 'lm', color = 'red', size = 2) + 
  theme_bw()
```
**Comment**: There is a strong positive relationship between `y` and `z`

4. Install the **ggplot2movies** package and look at the movies that have a
missing budget. How are they different from the movies with a budget?
```{r}
library(ggplot2movies)
data(movies)
```
First, let's get a look at the data and its structure
```{r}
movies
```


```{r}
movies %>% glimpse()
```
```{r}
# Create a new column `budget_type` (factor):
#   - 'missing' if `budget` is NA
#   - 'avaible' otherwise
mv <- movies %>% 
  mutate(budget_type = ifelse(is.na(budget), 'missing', 'avaible') %>% as_factor)

mv
```
```{r}
mv %>% 
  ggplot(aes(budget_type)) +
  geom_bar(fill = 'grey50') + 
  labs(x = 'budget type') + 
  theme_minimal() 

```
From the above plot, we can clearly see that the information about budget of most movies is missing

```{r}
# compare the distribution of 'missing budget' and 'available budget'
show_dist <- function(data, variable) 
  data %>% ggplot(aes({{variable}}, color = budget_type)) + geom_freqpoly() + theme_bw()

mv %>% show_dist(length)

mv %>% show_dist(rating)

```
5. What is `NA & FALSE` and `NA | TRUE`? Why? Why doesn’t `NA * 0` equal zero?
What number times zero does not equal 0? What do you expect `NA ˆ 0` to
equal? Why?

- `NA & FALSE` is `FALSE`. `NA` in this context is a missing boolean value. Any boolean **and** `FALSE` will return `FALSE`.

- `NA | TRUE` is `TRUE` (For the same reason as above)

- `NA * 0` is not zero because if `NA` is Infinity, then **Infinity multiply  0** does not make sense

- `NA ^ 0` is 1. In R, any number raise to power of 0 will always return 1

#  10.3 Create new vairables
See `?mutate`

To better explore the relationship between `x` and `y`, it’s useful to “rotate” the
plot so that the data is flat, not diagonal. We can do that by creating two new
variables: 
- one that represents the difference between x and y (which in this context represents the symmetry of the diamond) 
- one that represents its size (the length of the diagonal)

```{r}
# Remove outliers
diamonds_ok <- diamonds %>% filter(x > 0, y > 0, y < 20)

d <- diamonds_ok %>% 
  mutate(
    sym = x - y,
    size = sqrt(x * x + y * y)
  ) 

d %>%
  ggplot(aes(size, sym)) + 
  geom_bin2d() +
  theme_light()
```
This plot has two advantages: we can more easily see the pattern followed
by most diamonds, and we can easily select outliers. Here, it doesn’t seem
important whether the outliers are positive (i.e. x is bigger than y) or negative
(i.e. y is bigger x). So we can use the absolute value of the symmetry variable
to pull out the outliers. Based on the plot, and a little experimentation, I came
up with a threshold of 0.20

```{r}
d %>% ggplot(aes(abs(sym))) + geom_histogram(binwidth = .1) |
d %>% filter(abs(sym) < .2) %>% ggplot(aes(abs(sym))) + geom_histogram(binwidth = .01) 
```
**Comments**: While most diamonds are closed to be symmetric, there are very few that are perfectly symmetric (i.e: x == y)

## 10.3.1 Useful tools

Typically, transformations will be suggested by your domain knowledge. However, there are a few transformations that are useful in a surprisingly wide range of circumstances.

• Log-transformations are often useful. They turn multiplicative relationships into additive relationships; they compress data that varies over orders of magnitude; they convert power relationships to linear relationship.

• Relative difference: If you’re interested in the relative difference between two variables, use log(x / y). It’s better than x/y because it’s symmetric: if x < y, x/y takes values [0, 1), but if x > y, x/y takes values (1, Inf).

• Sometimes integrating or differentiating might make the data more interpretable: if you have distance and time, would speed or acceleration
be more useful? (or vice versa). (Note that integration makes data more
smooth; differentiation makes it less smooth.)

• Partition a number into magnitude and direction with `abs(x)` and `sign(x)`.
There are also a few useful ways to transform pairs of variables:
• Partitioning into overall size and difference is often useful, as seen above.

• If you see a strong trend, use a model to partition it into pattern and
residuals is often useful. You’ll learn more about that in the next chapter.

• Sometimes it’s useful to change positions to polar coordinates (or vice
versa): distance `(sqrt(xˆ2 + yˆ2))` and angle `(atan2(y, x))`.

## 10.3.2 Exercises
1. Practice your variable creation skills by creating the following new variables:
• The approximate volume of the diamond (using x, y, and z).
• The approximate density of the diamond.
• The price per carat.
• Log transformation of carat and price

```{r}
diamonds %>% 
  transmute(
    volume = 1 / 3 * x * y * depth,                   
    density = carat / volume, 
    ppc = price / carat ,         # price per carat
    log_price = log(price),
    log_carat = log(carat)
  )
```

2. How can you improve the data density of `ggplot(diamonds, aes(x, z)) + stat_bin2d()`.  
What transformation makes it easier to extract outliers?

```{r}
diamonds %>% 
  filter(x > 0, z > 0, z < 10) %>%                     # Remove obvious outliers
  mutate(sym = x - z, size = sqrt(x * x + z * z)) %>%  # Create new variables (symmetric and the the size)
  ggplot(aes(size, sym)) + 
  geom_bin2d() +
  theme_bw()
```
3. The depth variable is just the width of the diamond (average of x and y)
divided by its height (z) multiplied by 100 and round to the nearest integer.
Compute the depth yourself and compare it to the existing depth variable.
Summarise your findings with a plot.

4. Compare the distribution of symmetry for diamonds with `x > y` and `x < y`

```{r}
d %>% 
  filter(abs(sym) < .2) %>% 
  ggplot(aes(abs(sym), color = x > y)) + 
  geom_freqpoly(binwidth = .01) + 
  labs(x = 'symmetric', color = NULL) +
  scale_color_discrete(breaks = c(F, T), labels = c('x < y', 'x > y')) + 
  theme_light()
```
#  10.4 Group-wise summaries

1. Define the grouping variables with `group_by()`.
2. Describe how to summarise each group with a single row with `summarise()`

For example, to look at the average price per clarity, we first group by clarity
```{r message = F} 
price_clarity <- diamonds %>% group_by(clarity) %>% summarize(avg_price = mean(price, na.rm = T)) 

price_clarity
```
```{r}
price_clarity %>% 
  ggplot(aes(clarity, avg_price)) +
  geom_point(size = 4, color = 'tomato') + 
  geom_line(group = 1, color = 'grey80') +
  theme_dark()
```
You might be surprised by this pattern: WHy do diamonds with better clarity have lower prices?
We will be answering this question later

---
The next example shows how cut and depth interact
```{r}
diamonds %>% 
  ggplot(aes(depth, ..density.., color = cut)) +
  geom_freqpoly(binwidth = .1) + 
  scale_x_continuous(limits = c(55, 70)) +
  scale_color_brewer(palette = 'Dark2') + 
  theme_bw()
```
We can recreate the above plot by using `dplyr`:
```{r}
diamonds %>% 
  filter(depth %>% between(55, 70)) %>% 
  count(cut, depth) %>%                            # The frequency of each (cut, depth)
  group_by(cut) %>%mutate(prop = n / sum(n)) %>%   # Normalize the distribution of `depth` for each `cut`
  ggplot(aes(depth, prop, color = cut)) +
  geom_line() +
  theme_light() + 
  scale_color_brewer(palette = 'Dark2')
```

## 10.4.1 Useful tools

`summarise()` needs to be used with functions that take a vector of n values
and always return a single value. Those functions include:
• Counts: `n()`, `n_distinct(x)`.
• Middle: `mean(x)`, `median(x)`.
• Spread: `sd(x`), `mad(x)`, `IQR(x)`.
• Extremes: `quartile(x)`, `min(x)`, `max(x)`.
• Positions: `first(x)`, `last(x)`, `nth(x, 2)`

Most summary functions have a `na.rm` argument: `na.rm = TRUE` tells the
summary function to remove any missing values prior to summarisation. This
is a convenient shortcut: rather than removing the missing values then summarising, you can do it in one step.

## 10.4.2 Statistical Considerations
When summarising with the mean or median, it’s always a good idea to
include a count and a measure of spread. This helps you calibrate your
assessments—if you don’t include them you’re likely to think that the data is
less variable than it really is, and potentially draw unwarranted conclusions.


The following example summarizes of the average price
by clarity to also include the number of observations in each group, and the
upper and lower quartiles. It suggests the mean might be a bad summary for
this data - the distributions of price are so highly skewed that the mean is
higher than the upper quartile for some of the groups


```{r message = F}
diamonds %>% 
  group_by(clarity) %>% 
  summarize(avg_price= mean(price), 
            n = n(), 
            lq = quantile(price, .25),
            uq = quantile(price, .75))
```

```{r}
diamonds %>% 
  ggplot(aes(clarity, price)) +
  geom_boxplot(varwidth = T) +
  theme_minimal()
```
Another example of this comes from baseball. Let’s take the MLB batting
data from the Lahman package and calculate the batting average: the number
of hits divided by the number of at bats. Who’s the best batter according to
this metric?

```{r}
data(Batting, package = 'Lahman')
batters <- Batting %>% filter(AB > 0, !is.na(H), !is.na(AB))

batters
```

```{r message = F}
per_player <- batters %>% group_by(playerID) %>% summarize(success_rate = sum(H) / sum(AB))

per_player %>% ggplot(aes(success_rate)) + 
  geom_histogram(binwidth = .01) + 
  xlab('success rate') + 
  theme_minimal()
```
Wow, there are a lot of players who can hit the ball every single time (i.e: success_rate = 1). Why? Are they superman?
Let's double check

```{r message = F}
ba <- batters %>% group_by(playerID) %>% summarize(success_rate = sum(H) / sum(AB), total_at_bats = sum(AB))

ba %>% ggplot(aes(total_at_bats, success_rate)) +
  geom_bin2d(bins = 100) +
  geom_smooth() + 
  labs(x = 'Total number of at bats (AB)', y = 'Success Rate') +
  theme_minimal()
```
The highest batting averages occur for the players with the smallest number of at bats - it’s not hard to hit the ball every time if you’ve only had two pitches. We can make the pattern a little more clear by getting rid of the
players with less than 10 at bats.

```{r}

ba %>% filter(total_at_bats > 10) %>%  
  ggplot(aes(total_at_bats, success_rate)) +
  geom_bin2d(bins = 100) +
  geom_smooth() +
  theme_minimal()

```

## 10.4.3 Exercises

1. For each year in the `ggplot2movies::movies` data determine the percent of movies with missing budgets. Visualise the result.


```{r}
movies
```

```{r}
movies %>% 
  group_by(year) %>% 
  summarize(missing_rate = sum(is.na(budget)) / n()) %>%
  ggplot(aes(year, missing_rate)) +
  geom_line(color = 'tomato') +
  labs(x ='Year', y = 'Missing rate') + 
  theme_bw()
```
2. How does the average length of a movie change over time? Display your
answer with a plot, including some display of uncertainty

The average length of a movie tends to increase overtime. 

```{r message = F}
# Calculate Confidence Interval using Rmisc
library(Rmisc)

movies %>% 
  filter(!is.na(length)) %>%
  group_by(year) %>%
  # 95% Confidence Interval, NOTE: have to use `list` function, each cell is a named list: $lower, $mean, $upper
  transmute(ci = list(CI(length, .95))) %>% 
  # Unnest named list from each cell to wider format
  unnest_wider(ci) %>%
  ggplot(aes(year, mean, ymin = lower, ymax = upper)) + 
  geom_ribbon(alpha = .4) + 
  geom_line(color = 'tomato', size = 1) +
  ylab('Averge length time') + 
  theme_bw()
```
4. Compute a histogram of carat by “hand” using a binwidth of 0.1. Display
the results with geom bar(stat = "identity"). (Hint: you might need to
create a new variable first.)

```{r}
carat_hist <- diamonds %>% 
  pull(carat) %>%
  cut_width(width = .1) %>% 
  fct_count()

carat_hist
```

```{r}
carat_hist %>% 
  ggplot(aes(f, n)) +
  geom_col()
```
```{r}
diamonds %>% ggplot(aes(carat)) + geom_histogram(binwidth = .1)
```
5. In the baseball example, the batting average seems to increase as the
number of at bats increases. Why?

# 10.5 Transformation Pipelines
Using `%>%`

##  10.5.1 Exercises