---
title: "Data Analysis"
author: "Pikachu"
date: "11/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE} 
library(tidyverse)
library(patchwork)
```

## 9.3.3 Exercises

2. How can you convert back and forth between the `economics` and
`economics long` datasets built into ggplot2?

```{r}
# from long to wide
 economics_long %>% 
  mutate(value = value + value01, value01 = NULL) %>% 
  pivot_wider(names_from = variable, values_from = value)
```

```{r}
# from long to wide
economics %>% 
  pivot_longer(!date, names_to = 'variable', values_to = 'val') %>%
  mutate(value = trunc(val), value01 = val - value) %>%              # Separate the integer and float point part
  mutate(val = NULL) %>%                                             # Remove column `val`
  arrange(variable, date)
```

# 9.5 Case studies

## 9.5.1 Blood pressure

The first step when tidying a new dataset is always to identify the variables.
Look at the following simulated dataset.  
There are 6 variables:  name, age, start date, week, systolic & diastolic blood pressure.
```{r}
bpd <- readr::read_table(
"name age      start  week1  week2  week3
Anne  35 2014-03-27 100/80 100/75 120/90
Ben   41 2014-03-09 110/65 100/65 135/70
Carl  33 2014-04-02 125/80  <NA>   <NA>
", na = "<NA>")

bpd
```
```{r}
# Tidy dataset
bpd_1 <- bpd %>% pivot_longer(
  starts_with('week'),
  names_prefix = 'week',
  names_to = 'week',
  values_to = 'bp',
  names_transform = list(week = as.integer)
  ) %>% separate('bp', c('sys', 'dias'), convert = T) %>% 
  arrange(name, week)

bpd_1
```

You might notice that there’s some repetition in this dataset: if you know
the name, then you also know the age and start date. This reflects a third
condition of tidyness that I don’t discuss here: each data frame should contain
one and only one data set. Here there are really two datasets: information
about each person that doesn’t change over time, and their weekly blood
pressure measurements. 

```{r}
# assign a unqiue id for each patient
bpd_group <- bpd_1 %>% group_by(name, age, start) %>% mutate(id = cur_group_id()) %>% ungroup()

bpd_group
```

```{r}
# Information about each person
information <- bpd_group %>% distinct(id, name, age, start)
information
```

```{r}
# Weekly blood pressure
weekly_bp <- bpd_group %>% select(id, sys, dias)
weekly_bp
```
```{r}
# You can restore the original dataset by joining 2 dataset

information %>% right_join(weekly_bp, by = 'id') %>%
  mutate(id = NULL) # remove column id

```

## 9.5.2 Test scores
Imagine you’re interested in the effect of an intervention on test scores. You’ve
collected the following data. What are the variables?

```{r}
scores <- dplyr::data_frame(
person = rep(c("Greg", "Sally", "Sue"), each = 2),
time = rep(c("pre", "post"), 3),
test1 = round(rnorm(6, mean = 80, sd = 4), 0),
test2 = round(jitter(test1, 15), 0)
)

scores
```
There are 4 variables: person name, test name, pre score and post score

```{r}
scores_tidy <- scores %>% 
  pivot_longer(
    starts_with('test'),
    names_to = 'test',
    values_to = 'score',
    names_transform = list(test = parse_factor)
  ) %>%
  pivot_wider(
    names_from = 'time', 
    values_from = 'score',
    names_glue = '{time}_score'
  )

scores_tidy
```


A good indication that we have made a tidy dataset is that it’s now easy
to calculate the statistic of interest: the difference between pre- and postintervention scores:
```{r}
scores_tidy %>% mutate(diff = post_score - pre_score)
```

And it’s similarly easy to plot:

```{r}
scores_tidy %>% 
  mutate(diff = post_score - pre_score) %>%
  ggplot(aes(person, diff)) +
  geom_path(aes(group = person), arrow = arrow(15, unit(.4, 'cm')), size = .5) +
  geom_point(aes(color = test), size = 3) +
  theme_light()
```

