---
title: "Layer"
author: "Pikachu"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
library(dslabs)
library(titanic)
```

# 5.1 Introduction
# 5.2 Building a plot
# 5.3 Data
Every layer must have some data associated with it, and that data must be
in a tidy data frame.

The data on each layer doesn’t need to be the same and it’s often useful
to combine multiple datasets in a single plot.

```{r}

ggplot(economics) + 
  geom_line(aes(date, unemploy)) +
  geom_rect(data = presidential,
            mapping = aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf, fill = party),
            alpha = .5) +
  geom_vline(data = presidential,
             mapping = aes(xintercept = start),
             color = 'grey50') + 
  theme_light()
```


## 5.3.1 Exercises
1. The first two arguments to ggplot are `data` and `mapping`. The first two
arguments to all layer functions are `mapping` and `data`. Why does the order
of the arguments differ? (Hint: think about what you set most commonly.)

2. The following code uses dplyr to generate some summary statistics about
each class of car '


```{r}
mpg %>%
  ggplot(aes(class, hwy)) +
  geom_jitter(width = .2) +
  stat_summary(fun = mean, geom = 'point', size = 5,  color = 'tomato') + 
  stat_summary(fun = dplyr::n, geom = 'text') + 
  theme_light() 
```

# 5.4 Aesthetic mapping

Using `aes()`

```{r}
mpg %>%
  ggplot(aes(displ, cty, color = class)) + 
  geom_point() + 
  theme_bw()
```

# 5.4.1 Specifying the Aesthetics in the Plot vs. in the Layers

Aesthetic mappings can be supplied in the initial ggplot() call, in individual
layers, or in some combination of both. All of these calls create the same plot
specification:
```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) +
geom_point()
ggplot(mpg, aes(displ, hwy)) +
geom_point(aes(colour = class))
ggplot(mpg, aes(displ)) +
geom_point(aes(y = hwy, colour = class))
ggplot(mpg) +
geom_point(aes(displ, hwy, colour = class))
```
Within each layer, you can add, override, or remove mappings:

* add
```{r}
mpg %>%
  ggplot(aes(displ, hwy)) + 
  geom_point(aes(color = class))

#Equivalent to: mpg %>% ggplot(aes(displ, hwy, color = class)) + geom_point()
```
* Override

```{r}
mpg %>%
  ggplot(aes(displ, hwy, color = class)) + 
  geom_point(aes(color = manufacturer))

#Equivalent to: mpg %>% ggplot(aes(displ, hwy, color = manufacturer)) + geom_point()
```

* Remove
```{r}
mpg %>%
  ggplot(aes(displ, hwy, color = class)) + 
  geom_point(aes(color = NULL))

#Equivalent to: mpg %>% ggplot(aes(displ, hwy)) + geom_point()
```

## 5.4.2 Setting vs Mapping
If you want appearance to be governed by a variable, put the specification
inside `aes()`; if you want override the default size or colour, put the value
outside of `aes()`.
```{r}
#Mapping: This will add an legend

mpg %>%
  ggplot(aes(displ, hwy, color = 'tomato')) + 
  geom_point() |
  
#Setting 
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point(color = 'tomato')
```
It’s sometimes useful to map aesthetics to constants. For example, if you
want to display multiple layers with varying parameters, you can “name”
each layer:

```{r}
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() + 
  geom_smooth(aes(color = 'lm'), method = 'lm') + 
  geom_smooth(aes(color = 'loess'), method = 'loess') +
  labs(color = 'Method') 
```
## 5.4.3 Exercises

1. Simplify the following plot specifications:

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point()

ggplot(mpg, aes(cty, hwy)) + 
  geom_smooth()

ggplot(msleep, aes(brainwt, bodywt)) +
  geom_point() +
  scale_x_log10() + 
  scale_y_log10()
```
2. What does the following code do? Does it work? Does it make sense?
Why/why not?

```{r}
ggplot(mpg) +
geom_point(aes(class, cty)) +
geom_boxplot(aes(trans, hwy)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
3. What happens if you try to use a continuous variable on the x axis in one
layer, and a categorical variable in another layer? What happens if you do
it in the opposite order?

```{r}
mpg %>%
  ggplot(aes(displ, class)) + 
  geom_jitter(height = .2) + 
  theme_light()

mpg %>%
  ggplot(aes(class, displ)) + 
  geom_jitter(width = .2) + 
  theme_light()
```

# 5.5 Geoms
## 5.5.1 Exercises
1. Download and print out the ggplot2 cheatsheet from ![](http://www.rstudio.com/resources/cheatsheets/) so you have a handy visual reference for all the geoms.

2. Look at the documentation for the graphical primitive geoms. Which aesthetics do they use? How can you summarise them in a compact form?
Line: 
  * color: The color of the line
  * size: The linewidth
  * linetype: linestyle
Point: 
  * size: The size of the point
  * shape: The shape of the point (marker) - diamond, rectangle, circle, plus, square, ...
  * fill: The face color (For shape that can be filled, e.g: 19, 20, 21, ...)
  * stroke: The marker edge width (for shape that can be filled and have edge)
  * linetype: The line style of edge
  * color: edge color
Polygon:
 * fill: Facecolor
 * color: edge color
 * size: the linewidth of the edge
 * linetype: line style of the edge

3. What’s the best way to master an unfamiliar geom? List three resources
to help you get started.

- Read the documentation, go through all examples
- resources: ggplot2.tidyverse.com
- GGPLOT2 cheat sheet
4. For each of the plots below, identify the geom used to draw it.

```{r}
mpg %>%
  ggplot(aes(drv, displ)) + 
  geom_violin()

mpg %>%
  ggplot(aes(hwy, cty, size = displ)) + 
  geom_point(alpha = .3)

mpg %>%
  ggplot(aes(hwy, cty)) +
  geom_hex(binwidth = 3)

mpg %>%
  ggplot(aes(cyl, drv)) + 
  geom_jitter(width = .1, height = .1)

economics %>%
  ggplot(aes(date, psavert)) + 
  geom_area()

economics %>%
  ggplot(aes(uempmed, psavert, color = date)) + 
  geom_path()


```
5. For each of the following problems, suggest a useful geom:
• Display how a variable has changed over time: `geom_line`, `geom_bar`, `geom_col`
• Show the detailed distribution of a single variable. `geom_historgam`, `geom_density`, `geom_freqpoly`
• Focus attention on the overall trend in a large dataset. `geom_line`
• Draw a map. `geom_polygon`,
• Label outlying points. `geom_text`

# 5.6 Stats

There are two ways to use these functions. You can either add a `stat_()`
function and override the default geom, or add a `geom_()` function and override
the default stat:

```{r}
#Survival rate in each Pclass
titanic_train %>%
  ggplot(aes(Pclass, Survived)) +
  stat_summary(fun = mean, geom = 'bar')

#Equivalent
titanic_train %>%
  ggplot(aes(Pclass, Survived)) +
  geom_bar(stat = 'summary', fun = mean)
```

## 5.6.1 Generated variables

Internally, a stat takes a data frame as input and returns a data frame as
output, and so a stat can add new variables to the original dataset. It is
possible to map aesthetics to these new variables. For example, stat bin, the
statistic used to make histograms, produces the following variables:
• *count*, the number of observations in each bin
• *density*, the density of observations in each bin (percentage of total/bar
width)
• *x*, the centre of the bin

These generated variables can be used instead of the variables present in
the original dataset.

To refer to a
generated variable like density, “..” must surround the name: e.g: ..count..

```{r}
#Using a density histogram instead of a frequency hisogram
mpg %>%
  ggplot(aes(x = displ, y = ..density..)) + 
  geom_histogram(binwidth = .25, color = 'grey80')
```
This technique is particularly useful when you want to compare the distribution of multiple groups that have very different sizes. For example, it’s
hard to compare the distribution of `price` within `cut` because some groups
are quite small. It’s easier to compare if we standardise each group to take
up the same area:


```{r}
diamonds %>%
  ggplot(aes(price, color = cut)) + 
  geom_freqpoly(binwidth = 500, size = 1, show.legend = F) +
  theme_bw() |
diamonds %>%
  ggplot(aes(price, ..density.., color = cut)) + 
  geom_freqpoly(binwidth = 500, size = 1, show.legend = F) + 
  theme_bw()
```
The result of this plot is rather surprising: low quality diamonds seem to
be more expensive on average.

## 5.6.2 Exercises
1. The code below creates a similar dataset to `stat_smooth()`. Use the appropriate geoms to mimic the default `geom_smooth()` display.
```{r}
#From scratch
mod <- loess(hwy ~ displ, data = mpg)
smoothed <- data.frame(displ = seq(1.6, 7, length = 50))
pred <- predict(mod, newdata = smoothed, se = TRUE)
smoothed$hwy <- pred$fit
smoothed$hwy_lwr <- pred$fit - 1.96 * pred$se.fit
smoothed$hwy_upr <- pred$fit + 1.96 * pred$se.fit

smoothed %>%
  ggplot(aes(x = displ, y = hwy, ymin = hwy_lwr, ymax = hwy_upr)) + 
  geom_line(color = 'blue', size = 1) +
  geom_ribbon(fill = 'grey50', alpha = .3)

#ggplot2 function
mpg %>%
  ggplot(aes(displ, hwy)) + 
  geom_smooth()
```
2. What stats were used to create the following plots?
```{r}
mpg %>%
  mutate(mean_displ = mean(displ), sd_displ = sd(displ)) %>%
  ggplot(aes(sample = displ)) +
  geom_qq()

mpg %>%
  ggplot(aes(displ)) + 
  geom_density()
```

3. Read the help for stat sum() then use geom count() to create a plot that
shows the proportion of cars that have each combination of drv and trans.

```{r}
mpg %>%
  ggplot(aes(drv, trans, size = ..prop.., group = 1)) + 
  geom_count()
```
# 5.7 Position Adjustment
Position adjustments apply minor tweaks to the position of elements within
a layer. Three adjustments apply primarily to bars:

• position_stack(): stack overlapping bars (or areas) on top of each other.
• position_fill(): stack overlapping bars, scaling so the top is always at 1.
• position_dodge(): place overlapping bars (or boxplots) side-by-side.
• position_identity(): Do nothing

```{r}
diamonds %>%
  ggplot(aes(x = color, color = cut)) +
  geom_line(aes(group = cut), stat = 'count')
```
## 5.7.1 Exercises
1. When might you use `position_nudge()`? Read the documentation.
For side by side bar

```{r}
titanic_train %>%
  ggplot(aes(factor(Survived), fill = factor(Pclass))) +
  geom_bar(position = 'dodge')
```

2. Many position adjustments can only be used with a few geoms. For example, you can’t stack boxplots or errors bars. Why not? What properties
must a geom possess in order to be stackable? What properties must it
possess to be dodgeable?
3. Why might you use geom jitter() instead of geom count()? What are the
advantages and disadvantages of each technique?
4. When might you use a stacked area plot? What are the advantages and
disadvantages compared to a line plot?
