---
title: "Positioning"
author: "Pikachu"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 7.1 Introduction

* Position adjustment: position_stack(), position_dodge(), ...
* Position scales: scale_x_continuous(breaks, labels), ...
* Facetting
* Coordinate system

# 7.2 Facetting
## 7.2.1 Facet wrap

```{r}
iris %>%
  ggplot(aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point() + 
  facet_wrap(~Species, ncol = 2)
```

## 7.2.2  Facet grid

```{r}
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() + 
  facet_grid(class ~ cyl)
```

## 7.2.3 Controlling scale

• scales = "fixed": x and y scales are fixed across all panels.
• scales = "free x": the x scale is free, and the y scale is fixed.
• scales = "free y": the y scale is free, and the x scale is fixed.
• scales = "free": x and y scales vary across panels.

Fixed scales make it easier to see patterns across panels; free scales make
it easier to see patterns within panels.

```{r}
mpg2 %>%
  ggplot(aes(cty, model)) +
  geom_point() + 
  facet_grid(manufacturer ~ ., scales = 'free', space = 'free') +
  theme(strip.text.y = element_text(angle = 0))
```
## 7.2.4 Missing Facetting Variables

If you are using facetting on a plot with multiple datasets, what happens
when one of those datasets is missing the facetting variables?

```{r}
#remove column species from `iris`
iris1 <- iris %>% mutate(Species = NULL)
```

Here ggplot will do what you expect:
it will display the map in every facet: missing facetting variables are treated
like they have all values.


```{r}
ggplot(iris, aes(Sepal.Length, Sepal.Width)) + 
 
  geom_point(data = iris1, color = 'grey80') + 
  geom_point(aes(color = Species)) +         #Display original data in every facet
  facet_wrap(~ Species)
```
## 7.2.5 Grouping vs Faceting
Facetting is an alternative to using aesthetics (like colour, shape or size) to
differentiate groups. Both techniques have strengths and weaknesses, based
around the relative positions of the subsets. With facetting, each group is
quite far apart in its own panel, and there is no overlap between the groups.
This is good if the groups overlap a lot, but it does make small differences
harder to see. When using aesthetics to differentiate groups, the groups are
close together and may overlap, but small differences are easier to see.


## 7.2.6 Continous Variables

To facet continuous variables, you must first discretise them. ggplot2 provides
three helper functions to do so:
• Divide the data into n bins each of the same length: `cut interval(x, n)`
• Divide the data into bins of width width: `cut width(x, width)`.
• Divide the data into n bins each containing (approximately) the same
number of points: `cut number(x, n = 10)`.

## 7.2.7 Exercises
1. Diamonds: display the distribution of price conditional on cut and carat.
Try facetting by cut and grouping by carat. Try facetting by carat and
grouping by cut. Which do you prefer?

```{r}
diamonds %>%
  ggplot(aes(price, group = carat)) +
  geom_histogram(binwidth = 250) +
  facet_wrap(~ cut)
```

2. Diamonds: compare the relationship between price and carat for each
colour. What makes it hard to compare the groups? Is grouping better
or facetting? If you use facetting, what annotation might you add to make
it easier to see the differences between panels?
```{r}
diamonds %>%
  ggplot(aes(carat, price, color = color)) +
  geom_point()

diamonds %>% 
  ggplot(aes(carat, price)) + 
  geom_point(alpha = .2) + 
  facet_wrap(~ color)
```

3. Why is `facet wrap()` generally more useful than `facet grid()`?

4. Recreate the following plot. It facets mpg2 by class, overlaying a smooth
curve fit to the full dataset.

```{r}
mpg2 %>%
  ggplot(aes(displ, hwy)) + 
  geom_point() +
  geom_smooth(data = mpg2 %>% mutate(class = NULL)) + 
  facet_wrap(~ class) 
```
# 7.3 Coordinate System

There are two types of coordinate system. Linear coordinate systems preserve the shape of geoms:
• coord cartesian(): the default Cartesian coordinate system, where the 2d
position of an element is given by the combination of the x and y positions.
• coord flip(): Cartesian coordinate system with x and y axes flipped.
• coord fixed(): Cartesian coordinate system with a fixed aspect ratio.
On the other hand, non-linear coordinate systems can change the shapes:
a straight line may no longer be straight. The closest distance between two
points may no longer be a straight line.
• coord map()/coord quickmap(): Map projections.
• coord polar(): Polar coordinates.
• coord trans(): Apply arbitrary transformations to x and y positions, after
the data has been processed by the stat

# 7.4 Linear Coordinate System

There are three linear coordinate systems: `coord cartesian()`, `coord flip()`,
`coord fixed()`.

## 7.4.1 Zooming into a Plot with `coord_cartesian()`

## 7.4.2 Flipping the Axes with `coord_flip()`

## 7.4.3 Equal scales with `coord_fixed()`

# 7.5 Non-linear Coordinate System

## 7.5.1 Transformations with `coord_trans()`


Transforming at the
coordinate system level occurs after the statistics have been computed, and
does affect the shape of the geom.

## 7.5.2 Polar Coordinates with `coord_polar()`

## 7.5.3 Map Projections with `coord_map()`

