---
title: "Programming with ggplot2"
author: "Pikachu"
date: "11/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
```

# 12.1 Introduction
A major requirement of a good data analysis is flexibility. If your data
changes, or you discover something that makes you rethink your basic assumptions, you need to be able to easily change many plots at once. The main inhibitor of flexibility is code duplication. If you have the same plotting statement repeated over and over again, you’ll have to make the same change in many different places. 

To make your code more flexible, you need to reduce duplicated code bywriting functions. When you notice you’re doing the same thing over and overagain, think about how you might generalise it and turn it into a function.

In this chapter I’ll show how to write functions that create:
• A single ggplot2 component.
• Multiple ggplot2 components.
• A complete plot.

# 12.2 Single Components
Each component of a ggplot plot is an object. Most of the time you create the
component and immediately add it to a plot, but you don’t have to. Instead,
you can save any component to a variable (giving it a name), and then add
it to multiple plots:
```{r}
point <- geom_point(size = 3, shape = 21)

iris %>% ggplot(aes(Sepal.Length, Sepal.Width)) + point |
  
iris %>% ggplot(aes(Petal.Length, Petal.Width)) + point

```
That’s a great way to reduce simple types of duplication (it’s much better
than copying-and-pasting!), but requires that the component be exactly the
same each time. If you need more flexibility, you can wrap these reusable
snippets in a function.

```{r}
my_point <- function(size = 3, shape = 21, ...) geom_point(size = size, shape = shape, ...)

iris %>% ggplot(aes(Sepal.Length, Sepal.Width)) + my_point() |
  
iris %>% ggplot(aes(Petal.Length, Petal.Width)) + my_point(shape = 5, color = 'tomato')
```


Finally, note that you can only *add* components to a plot; you can’t modify
or remove existing objects.

## 12.2.1 Exercises
1. Create an object that represents a pink histogram with 100 bins.

```{r}
pink_hist <- geom_histogram(bins = 100, fill = 'pink')

diamonds %>% ggplot(aes(carat)) + pink_hist
```

2. Create an object that represents a fill scale with the Blues ColorBrewer
palette.

```{r}
blues <- scale_fill_brewer(palette = 'Blues')
```

3. Read the source code for `theme_grey()`. What are its arguments? How does
it work?

4. Create `scale_colour_wesanderson()`. It should have a parameter to pick the
palette from the `wesanderson` package, and create either a continuous or
discrete scale

```{r}
scale_color_wesanderson <- function(..., type = 'discrete', palette = 'Royal1', aesthetics = 'colour') {
  discrete_scale(aesthetics, 'wesanderson', function(n) wes_palette(type = type, name = palette), ...)
}

p <- iris %>% ggplot(aes(Sepal.Length, Sepal.Width, color = Species)) + geom_point(show.legend = F) 

p + scale_color_wesanderson() + theme_dark() | 

p + scale_color_wesanderson(palette = 'BottleRocket1') + theme_light()
```

# 12.3 Multiple Components

```{r message = F}
library(Hmisc)
geom_mean <- function() {
list(
stat_summary(fun = "mean", geom = "bar", fill = "grey70"),
stat_summary(fun.data = "mean_cl_normal", geom = "errorbar", width = 0.4)
)
}
ggplot(mpg, aes(class, cty)) + geom_mean()
ggplot(mpg, aes(drv, cty)) + geom_mean()
```


If the list contains any NULL elements, they’re ignored. This makes it easy
to conditionally add components:

```{r}
geom_mean <- function(se = TRUE) {
list(
stat_summary(fun = "mean", geom = "bar", fill = "grey70"),
if (se)
stat_summary(fun.data = "mean_cl_normal", geom = "errorbar", width = 0.4)
)
}
ggplot(mpg, aes(drv, cty)) + geom_mean()
ggplot(mpg, aes(drv, cty)) + geom_mean(se = FALSE)
```
## 12.3.1 Plot Components

You’re not just limited to adding layers in this way. You can also include any
of the following object types in the list:

• A data.frame, which will override the default dataset associated with the
plot. (If you add a data frame by itself, you’ll need to use %+%, but this is
not necessary if the data frame is in a list.)
• An aes() object, which will be combined with the existing default aesthetic
mapping.
• Scales, which override existing scales, with a warning if they’ve already
been set by the user.
• Coordinate systems and facetting specification, which override the existing
settings.
• Theme components, which override the specified components.

## 12.3.2 Annotations

It’s often useful to add standard annotations to a plot. In this case, your
function will also set the data in the layer function, rather than inheriting it
from the plot. There are two other options that you should set when you do
this. These ensure that the layer is self-contained:
• `inherit.aes = FALSE` prevents the layer from inheriting aesthetics from the
parent plot. This ensures your annotation works regardless of what else is
on the plot.
• `show.legend = FALSE` ensures that your annotation won’t appear in the
legend.

## 12.3.3 Additional Arguments

Read this section in the book

## 12.3.4 Exercises

# 12.4 Plot Functions

Instead of creating components, you might want to write a
function that takes data and parameters and returns a complete plot.
For example, you could wrap up the complete code needed to make a
piechart:


```{r}
piechart <- function(data, mapping) {
ggplot(data, mapping) +
geom_bar(width = 1) +
coord_polar(theta = "y") +
xlab(NULL) +
ylab(NULL)
}
piechart(mpg, aes(factor(1), fill = class))
```

## 12.4.1 Indirectly Referring to Variables

Read the book

```{r}
# Take the name of variable and show its distribution
show_dist <- function(data, var) data %>% ggplot(aes_(as.name(var))) + geom_histogram()

iris %>% show_dist('Sepal.Length')
```

## 12.4.2 The Plot Environment

Read the book

## 12.5 Functional Programming
Since ggplot2 objects are just regular R objects, you can put them in a list.
This means you can apply all of R’s great functional programming tools. For
example, if you wanted to add different geoms to the same base plot, you
could put them in a list and use `lapply()`.

```{r}
geoms <- list(
geom_point(),
geom_boxplot(aes(group = cut_width(displ, 1))),
list(geom_point(), geom_smooth())
)
p <- ggplot(mpg, aes(displ, hwy))
lapply(geoms, function(g) p + g)
```
## 12.5.1 Exercises
1. How could you add a `geom_point()` layer to each element of the following list?

```{r}
plots <- list(
ggplot(mpg, aes(displ, hwy)),
ggplot(diamonds, aes(carat, price)),
ggplot(faithfuld, aes(waiting, eruptions, size = density))
)
```

Solution
```{r}
plots %>% lapply(as_mapper(~ . + geom_point()))
```

