---
title: "Scales Axes Legend"
author: "Pikachu"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
library(dslabs)
```

# 6.1 Introduction
# 6.2 Scales
## 6.2.1 Exercises
1. What happens if you pair a discrete variable to a continuous scale? Raise an error

What
happens if you pair a continuous variable to a discrete scale?
```{r}
mpg %>%
  ggplot(aes(class, displ)) + 
  geom_jitter(width = .2) +
  scale_x_continuous()

mpg %>%
  ggplot(aes(cty, hwy)) + 
  geom_point() + 
  scale_x_discrete()
```

2. Simplify the following plot specifications to make them easier to understand.
```{r}
mpg %>%
  ggplot(aes(displ, hwy)) + 
  geom_point() +
  scale_y_continuous(name = 'Highway mpg')


mpg %>%
  ggplot(aes(class, displ, color = drv)) + 
  geom_point() +
  labs(x = 'Type of car', y = 'displayment', color = 'Drive train')

```
# 6.3 Guides: Axes and Legend

## 6.3.1 Scale title

`name`: Change the title of the scale. You can supply text or mathematical expression

```{r}
mpg %>%
  ggplot(aes(displ, hwy)) + 
  geom_point() +
  scale_x_continuous(name = quote(a + b ^ c)) +
  scale_y_continuous(name = 'Fuck\nyou')
```

Because tweaking these labels is such a common task, there are three
helpers that save you some typing: 
* `xlab()`
* `ylab()` 
* `labs()`

```{r}
mpg %>%
  ggplot(aes(displ, hwy, color = class)) + 
  geom_point() +
  xlab('displacement') +           #set x-axis label
  ylab('hightway mpg') +           #Set y-axis label
  labs(color = 'class of the car') #Set legend's title
```
Remove axis-label: Set to NULL
```{r}
mpg %>%
  ggplot(aes(displ, hwy, color = class)) + 
  geom_point() + 
  #Remove x-axis title, y-axis title and legend's title
  labs(x = NULL, y = NULL, color = NULL)
```
## 6.3.2 Breaks and labels

* breaks: set_xticks, set_yticks
* labels: set_xticklabels, set_yticklabels

```{r}
df <- data.frame(x = c(1, 3, 5) * 1000, y = 1)
axs <- ggplot(df, aes(x, y)) +
geom_point() +
labs(x = NULL, y = NULL)
axs
axs + scale_x_continuous(breaks = c(2000, 4000))
axs + scale_x_continuous(breaks = c(2000, 4000), labels = c("2k", "4k"))
```

```{r}
leg <- ggplot(df, aes(y, x, fill = x)) +
geom_tile() +
labs(x = NULL, y = NULL)
leg
leg + scale_fill_continuous(breaks = c(2000, 4000))
leg + scale_fill_continuous(breaks = c(2000, 4000), labels = c("2k", "4k"))
```

If you want to relabel the breaks in a categorical scale, you can use a
named labels vector:
```{r}
df2 <- data.frame(x = 1:3, y = c("a", "b", "c"))
ggplot(df2, aes(x, y)) +
geom_point()
ggplot(df2, aes(x, y)) +
geom_point() +
scale_y_discrete(labels = c(a = "apple", b = "banana", c = "carrot"))
```

To suppress breaks (and for axes, grid lines) or labels, set them to NULL:
```{r}
axs + scale_x_continuous(breaks = NULL)

axs + scale_x_continuous(labels = NULL)
```

Additionally, you can supply a function to breaks or labels. The breaks
function should have one argument, the limits (a numeric vector of length

two), and should return a numeric vector of breaks. The labels function
should accept a numeric vector of breaks and return a character vector of
labels (the same length as the input). The scales package provides a number
of useful labelling functions:

• scales::comma_format() adds commas to make it easier to read large numbers.
• scales::unit_format(unit, scale) adds a unit suffix, optionally scaling.
• scales::dollar_format(prefix, suffix) displays currency values, rounding
to two decimal places and adding a prefix or suffix.
• scales::wrap_format() wraps long labels into multiple lines.
See the documentation of the scales package for more details.


You can adjust the minor breaks (the faint grid lines that appear between the major grid lines) by supplying a numeric vector of positions to the
`minor_breaks` argument. This is particularly useful for log scales


## 6.3.3 Exercises
1. Recreate the following graphic:
```{r}
mpg %>%
  ggplot(aes(displ, hwy)) + 
  geom_point() +
  scale_x_continuous(labels = function(x) paste0(x, ' L')) +
  labs(x = 'Displacement', y = 'Highway mile per gallon')
```


2. List the three different types of object you can supply to the `breaks` argument. How do` breaks` and `labels` differ?

Supply to the  `breaks` argument:
* a list of values: e.g: c(10, 20, 100)
* a mapping: e.g: c('F' = 'Female', 'M' = 'Female')
* a function

3. Recreate the following plot:


```{r}
mpg %>%
  ggplot(aes(displ, hwy, color = drv)) + 
  geom_point() +
  scale_color_discrete(labels = function(v) paste0(v, 'wd'))
```
4. What label function allows you to create mathematical expressions? What
label function converts 1 to 1st, 2 to 2nd, and so on?

5. What are the three most important arguments that apply to both axes
and legends? What do they do? Compare and contrast their operation for
axes vs. legends.

* name
  - Axes: set axes title (x-axis title, y-axis title, ..)
  - Legend: set legend title
```{r}
#Axes title
mpg %>%
  ggplot(aes(displ, hwy)) + 
  geom_point() + 
  scale_x_continuous(name = 'displacement') + 
  scale_y_continuous(name = 'highway mpg') |

#Legend title
mpg %>%
  ggplot(aes(displ, hwy, color = class)) +
  geom_point() + 
  scale_color_discrete(name = 'class of car')
```

* breaks
* labels

```{r}
#Axes
mpg %>% ggplot(aes(displ, hwy)) + 
  geom_point() |
mpg %>% ggplot(aes(displ, hwy)) + 
  geom_point() + 
  scale_x_continuous(breaks = c(2, 5, 7), labels = c('Two', 'Five', 'Seven'))

#Legend
mpg %>%
  ggplot(aes(displ, hwy, size = cyl)) + 
  geom_point(alpha = .5) +
  theme_bw() |
mpg %>%
  ggplot(aes(displ, hwy, size = cyl)) + 
  geom_point(alpha = .5) +
  theme_bw() +
  scale_size_continuous(breaks = c(4, 7, 8), labels = c('Four', 'Seven', 'Eight'))
```
# 6.4 Legend
## 6.4.1 Layers and Legends

```{r}
df <- data.frame(x = 0, y = 1:3, z = letters[1:3])
```

By default, a layer will only appear if the corresponding aesthetic is
mapped to a variable with aes(). You can override whether or not a layer
appears in the legend with show.legend: FALSE to prevent a layer from ever
appearing in the legend; TRUE forces it to appear when it otherwise wouldn’t.
Using TRUE can be useful in conjunction with the following trick to make
points stand out:

```{r}
df %>%
  ggplot(aes(y, y)) + 
  geom_point(size = 4, color = 'grey20') + 
  geom_point(aes(color = z), size = 2) |

df %>%
  ggplot(aes(y, y)) + 
  geom_point(size = 4, color = 'grey20', show.legend = T) + 
  geom_point(aes(color = z), size = 2)
  
```
Sometimes you want the geoms in the legend to display differently to the
geoms in the plot.

```{r}
iris %>%
  ggplot(aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 3, alpha = .3) + 
  scale_color_discrete(guide = guide_legend(override.aes = list(alpha = 1)))
             
```
ggplot2 tries to use the fewest number of legends to accurately convey the
aesthetics used in the plot. It does this by combining legends where the same
variable is mapped to different aesthetics. The figure below shows how this
works for points: if both colour and shape are mapped to the same variable,
then only a single legend is necessary.

```{r}
ggplot(df, aes(x, y)) + geom_point(aes(colour = z)) |
ggplot(df, aes(x, y)) + geom_point(aes(shape = z)) |
ggplot(df, aes(x, y)) + geom_point(aes(shape = z, colour = z))
```
In order for legends to be merged, they must have the same `name`. So if you
change the name of one of the scales, you’ll need to change it for all of them.

## 6.4.2 Legend Layout

Position

```{r}
p <- mpg %>%
  ggplot(aes(displ, hwy, color = class)) + 
  geom_point() 

p

p + theme(legend.position = 'top')

p + theme(legend.position = 'none') #turn off legend
```
Position, horizontal alignment and vertical alignment

```{r}
p + 
  theme(legend.position = c(0, 1), legend.justification = c(0, 1)) #place legend at the top left of the panel

p + theme(legend.position = c(.5, .5), legend.justification = c(.5, .5))

p + theme(legend.position = c(1, 1), legend.justification  = c(1, 1))
```

# 6.4.3 Guide Function

# 6.4.4 Exercises
1. How do you make legends appear to the left of the plot?

Using  `theme(legend.position = 'left')`
```{r}
iris %>%
  ggplot(aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point() +
  theme(legend.position = 'left')
```

2. What’s gone wrong with this plot? How could you fix it?
multiple legends MUST have the same `name` to overlay each other
```{r}
#default
ggplot(mpg, aes(displ, hwy)) +
geom_point(aes(colour = drv, shape = drv)) +
scale_colour_discrete("Drive train") |

#Fixed
mpg %>%
  ggplot(aes(displ, hwy, color = drv, shape = drv)) + 
  geom_point() +
  scale_color_discrete('Same legend title') + 
  scale_shape_discrete('Same legend title')
```
3. Can you recreate the code for this plot?
```{r}
mpg %>%
  ggplot(aes(displ, hwy, color = class)) + 
  geom_smooth(method = 'lm', se = F) + 
  geom_point(show.legend = F) + 
  theme(legend.position = 'bottom',
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 1))
```
# 6.5 Limits

The limits, or domain, of a scale are usually derived from the range of the
data. There are two reasons you might want to specify limits rather than
relying on the data:
1. You want to make limits smaller than the range of the data to focus on
an interesting area of the plot.
2. You want to make the limits larger than the range of the data because you
want multiple plots to match up.


You can modify the limits using the limits parameter of the scale:
• For continuous scales, this should be a numeric vector of length two. If
you only want to set the upper or lower limit, you can set the other value
to NA.
• For discrete scales, this is a character vector which enumerates all possible
values.

```{r}
#default
p <- mpg %>% 
  ggplot(aes(class, displ)) + 
  geom_jitter(width = .2)

p |
  
#Continuous, 2 values
p + scale_y_continuous(limits = c(3, 5)) | #set y lim to range 3 -> 5

#Continuous, 1 value, set ylim starting from 4
p + scale_y_continuous(limits = c(4, NA)) | 
  
#Discete, only display  2seater and compact
p + scale_x_discrete(limits = c('2seater', 'compact'))
```

You can also use `xlim`, `ylim`, `lims`

```{r}
p |
p + ylim(3,5) |
p + ylim(4, NA) |
p + xlim(c('2seater', 'compact'))
```

If you have eagle eyes, you’ll have noticed that the range of the axes
actually extends a little bit past the limits that you’ve specified. This ensures
that the data does not overlap the axes. To eliminate this space, set `expand = c(0, 0)`. This is useful in conjunction with `geom_raster()`:

```{r}
f_plot <- faithfuld %>% 
  ggplot(aes(eruptions, waiting, fill = density)) + 
  geom_raster() + 
  theme(legend.position = 'none')

f_plot |
  
f_plot + 
  scale_x_continuous(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0,0))
```
## 6.5.1 Exercises

1. The following code creates two plots of the mpg dataset. Modify the code
so that the legend and axes match, without using facetting!
```{r}
#Default
fwd <- subset(mpg, drv == "f")
rwd <- subset(mpg, drv == "r")
ggplot(fwd, aes(displ, hwy, colour = class)) + 
  geom_point() |
ggplot(rwd, aes(displ, hwy, colour = class)) + 
  geom_point()

#Solution
fwd %>%
  ggplot(aes(displ, hwy, color = class)) + 
  geom_point() + 
  scale_x_continuous(limits = c(1, 7)) + 
  scale_y_continuous(limits = c(15, 45)) |

rwd %>%
  ggplot(aes(displ, hwy, color = class)) + 
  geom_point() + 
  xlim(c(1, 7)) + 
  ylim(c(15, 45))
```
```{r}
mpg %>%
  ggplot(aes(displ, hwy, color = class)) + 
  geom_point()
```
2. What does `expand_limits()` do and how does it work? Read the source
code.

3. What happens if you add two `xlim()` calls to the same plot? Why?
The last one will be executed, all previous will be ignored
```{r}
mpg %>%
  ggplot(aes(displ, hwy)) + 
  geom_point() + 
  xlim(3, 5) + 
  xlim(2, 6)
```
4. What does scale x continuous(limits = c(NA, NA)) do? think of it as range (-Inf, Inf)

```{r}
mpg %>%
  ggplot(aes(displ, hwy)) + 
  geom_point() + 
  scale_x_continuous(limits = c(NA, NA))
```

# 6.6 Scale tools

## 6.6.1 Continuous Position Scales

Every plot has two position scales, x and y. The most common continuous position scales are `scale x continuous()` and `scale y continuous()`, which
linearly map data to the x and y axis. The most interesting variations are
produced using transformations. Every continuous scale takes a `trans` argument, allowing the use of a variety of transformations:

```{r}
gapminder %>%
  filter(continent == 'Asia' & year == 2010 & !is.na(gdp)) %>% 
  mutate(country = reorder(country, gdp)) %>%
  ggplot(aes(country, gdp)) +
  geom_col() + 
  coord_flip() + 
  scale_y_continuous(trans = 'log10') +
  theme_bw()
```
# 6.7 Colors
