---
title: "Toolbox"
author: "Pikachu"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
```

# 3.1 Introduction

# 3.2 Basic plot types
## 3.2.1 Exercies

1. What geoms would you use to draw each of the following named plots?
* 1. Scatterplot: `geom_point()`
* 2. Line chart: `geom_line()`
* 3. Histogram: `geom_histogram()`
* 4. Bar chart: `geom_bar()`, `geom_col()`
* 5. Pie chart  `geom_bar(position = 'stack') + coord_polar(theta = 'y')` 

2. What’s the difference between `geom_path()` and `geom_polygon()`? What’s the
difference between geom `path()` and geom `line()`?

- Polygons are very similar to paths (as drawn by geom_path()) except that the start and end points are connected and the inside is coloured by fill.
- `geom_path()` connects the observations in the order in which they appear in the data. `geom_line()` connects them in order of the variable on the x axis.

3. What low-level geoms are used to draw geom `smooth()`? What about
`geom_boxplot()` and `geom_violin()`?

* `geom_smooth`: `geom_segment()`, `geom_ribborn()`
* `geom_boxplot()`: `geom_rect()`, `geom_segment()`
* `geom_violin()`: `geom_curve`

# 3.3 Labels
`geom text()`: which adds labels at the specified x and y positions

`geom_text()` has the most aesthetics of any geom, because there are so many
ways to control the appearance of a text:

* font-family: `family` argument
```{r}
family <- c('sans', 'serif', 'mono')

ggplot() + 
  annotate(y = 1:3, x = 0, label = family, geom = 'text', family = family, size = 10) +
  theme_void()
```
* font-face: `font-face` argument ('plain', 'bold', 'italic')
```{r}
face <- c('plain', 'bold', 'italic')
ggplot() + 
  annotate(y = 1:3, x = 0, label = face, geom = 'text', fontface = face) +
  theme_void()
```

* • You can adjust the alignment of the text with the `hjust` (“left”, “center”,
“right”, “inward”, “outward”) and `vjust` (“bottom”, “middle”, “top”, “inward”, “outward”) aesthetics.
The default alignment is centered. 
One of the most useful alignments is “inward”: it aligns text towards the middle
of the plot:
```{r}
df <- data.frame(
x = c(1, 1, 2, 2, 1.5),
y = c(1, 2, 1, 2, 1.5),
text = c(
"bottom-left", "bottom-right",
"top-left", "top-right", "center"
)
) 
ggplot(df, aes(x, y)) +
geom_text(aes(label = text)) | 
ggplot(df, aes(x, y)) +
geom_text(aes(label = text), vjust = "inward", hjust = "inward")
```

* `size` controls the font size. Unlike most tools, ggplot2 uses mm, rather
than the usual points (pts). This makes it consistent with other size units
in ggplot2. (There are 72.27 pts in a inch, so to convert from points to
mm, just multiply by 72.27/25.4.)

```{r}
sizes <- c(5, 10, 15)

ggplot() +
  annotate(x = 0, y = 1:3, label = paste0('size = ', sizes), size = sizes, geom = 'text') +
  theme_void()
```

*  `angle` specifies the rotation of the text in degrees.

```{r}
angles <- c(-135, -90, -45, 0, 45, 90, 135, 180)

ggplot() +
  annotate(y = 0, x = seq(angles), label = paste0('angle = ', angles), geom = 'text', angle = angles)
```
* Often you want to label existing points on the plot. You don’t want the
text to overlap with the points (or bars etc), so it’s useful to offset the text
a little. The `nudge_x` and `nudge_y` parameters allow you to nudge the text a
little horizontally or vertically:

```{r}
#Initially, The position of the text is (0, 0)
ggplot(data.frame(x = 0, y = 0), aes(x, y)) +
  geom_point() + 
  #nudge_x = 2, nudge_y = 3 will move text to (2, 3)
  geom_text(label = '(0, 0)', nudge_x = 2, nudge_y = 3, color = 'red')
```
* If `check_overlap = TRUE`, overlapping labels will be automatically removed.
The algorithm is simple: labels are plotted in the order they appear in the
data frame; if a label would overlap with an existing point, it’s omitted.
This is not incredibly useful, but can be handy.

```{r}
#overlapping labels
mpg %>%
  ggplot(aes(displ, cty, label = model)) + 
  geom_text(check_overlap = F) |
#remove overlapping labels
mpg %>%
  ggplot(aes(displ, cty, label = model)) +
  geom_text(check_overlap = T)
```
`geom_label()`: it draws a rounded rectangle
behind the text. This makes it useful for adding labels to plots with busy
backgrounds:

```{r}
label <- data.frame(
  waiting = c(55, 80),
  eruptions = c(2, 4.3),
  label = c("peak one", "peak two")
)
ggplot(faithfuld, aes(waiting, eruptions)) +
  geom_tile(aes(fill = density)) +
  geom_label(data = label, aes(label = label))
```
Labelling data well poses some challenges:

• Text does not affect the limits of the plot. Unfortunately there’s no way to
make this work since a label has an absolute size (e.g. 3 cm), regardless of
the size of the plot. This means that the limits of a plot would need to be
different depending on the size of the plot—there’s just no way to make
that happen with ggplot2. Instead, you’ll need to tweak `xlim()` and `ylim()`
based on your data and plot size.

• If you want to label many points, it is difficult to avoid overlaps.
`check_overlap = TRUE` is useful, but offers little control over which labels
are removed.




Text labels can also serve as an alternative to a legend. This usually makes
the plot easier to read because it puts the labels closer to the data. The
directlabels [directlabels](https://github.com/tdhock/directlabels) package, by Toby
Dylan Hocking, provides a number of tools to make this easier:

```{r}
library(directlabels)

mpg %>%
  ggplot(aes(displ, cty, color = class)) + 
  geom_point(show.legend = F) +                                     #turn off legend
  directlabels::geom_dl(aes(label = class), method = "smart.grid")  #plot direct label
```
# 3.4 Annotations

Annotations add metadata to your plot. But metadata is just data, so you
can use:

• `geom_text()` to add text descriptions or to label points Most plots will
not benefit from adding text to every single observation on the plot, but
labelling outliers and other important points is very useful.

• `geom_rect()` to highlight interesting rectangular regions of the plot.

• `geom_line()`, `geom_path()` and `geom_segment()` to add lines. All these geoms
have an arrow parameter, which allows you to place an arrowhead on the
line. Create arrowheads with `arrow()`, which has arguments *angle*, *length*,
*ends* and *type*

• `geom_vline()`, `geom_hline()` and `geom_abline()` allow you to add reference
lines (sometimes called rules), that span the full range of the plot..


To show off the basic idea, we’ll draw a time series of unemployment:

```{r}
economics %>%
  ggplot(aes(date, unemploy)) + 
  geom_line() +
  theme_classic()
```
We can annotate this plot with which president was in power at the time.

```{r}
ggplot(economics) +
  geom_line(aes(date, unemploy), size = 1) +
  geom_rect(data = presidential,
           mapping = aes(ymin = -Inf, ymax = Inf, xmin = start, xmax = end, fill = party),
           alpha = .5) +
  geom_text(data = presidential,
            mapping = aes(x = start, y = 2300, label = name),
            hjust = -.1, size = 3, vjust = 'inward') + 
  geom_vline(data = presidential, aes(xintercept = start), 
             color = 'white') + 
  scale_fill_brewer(palette = 'Set1') + 
  theme_minimal()
```
# 3.5 Collective Geoms

## 3.5.1 Multiple groups, one Aesthetic
The dataset we will be working with

```{r}
data(Oxboys, package = "nlme")
head(Oxboys)
```
In many situations, you want to separate your data into groups, but render
them in the same way. In other words, you want to be able to distinguish
individual subjects, but not identify them
```{r}
Oxboys %>%
  ggplot(aes(age, height, group = Subject)) + 
  geom_line() +
  geom_point() +
  theme_bw()
```
of multiple variables, use `interaction()` to combine them, e.g. `aes(group = interaction(school id, student id))`

## 3.5.2 Different groups on different layers

Add a single smooth line, showing the overall trend for all boys

```{r}
Oxboys %>%
  ggplot(aes(age, height)) + 
  #adding `group` in local layer instead of global layer (i.e: in ggplot(...))
  geom_line(aes(group = Subject)) +
  geom_point(aes(group = Subject)) +
  geom_smooth(method = 'lm', size = 2, se = F, color = 'red') +
  theme_bw()
```
## 3.5.3 Overriding the Default Grouping

Some plots have a discrete x scale, but you still want to draw lines connecting
across groups. 
Now we want to overlay lines that connect each individual boy

```{r}
Oxboys %>%
  ggplot(aes(Occasion, height)) + 
  geom_boxplot() + 
  geom_line(aes(group = Subject), color = 'blue') +
  theme_bw()
```

## 3.5.4 Matching Aesthetics to Graphic Objects

For bar and area plots, because stacking
the individual pieces produces the same shape as the original ungrouped data:

```{r}
mpg %>%
  ggplot(aes(class)) + 
  geom_bar() |
mpg %>%
  ggplot(aes(class, fill = drv)) + 
  geom_bar()
```
If you try to map fill to a continuous variable in the same way, it doesn’t
work. The default grouping will only be based on class, so each bar will be
given multiple colours. Since a bar can only display one colour, it will use the
default grey. To show multiple colours, we need multiple bars for each class,
which we can get by overriding the grouping:

```{r}
mpg %>%
  ggplot(aes(class, fill = hwy)) + 
  geom_bar() |
mpg %>%
  ggplot(aes(class, fill = hwy, group = hwy)) + 
  geom_bar()
```
The bars will be stacked in the order defined by the grouping variable. If
you need fine control, you’ll need to create a factor with levels ordered as
needed

## 3.5.5 Exercises
1. Draw a boxplot of `hwy` for each value of `cyl`, without turning `cyl` into a
factor. What extra aesthetic do you need to set?

set `group = cyl`
```{r}
mpg %>%
  ggplot(aes(cyl, hwy, group = cyl)) + 
  geom_boxplot()
```

2. Modify the following plot so that you get one boxplot per integer value of
`displ`
```{r}
ggplot(mpg, aes(displ, cty, group = displ)) +
  geom_boxplot()
```


3. When illustrating the difference between mapping continuous and discrete
colours to a line, the discrete example needed aes(group = 1). Why? What
happens if that is omitted? What’s the difference between aes(group = 1)
and aes(group = 2)? Why?


4. How many bars are in each of the following plots?
```{r}
ggplot(mpg, aes(drv)) +
geom_bar()

ggplot(mpg, aes(drv, fill = hwy, group = hwy)) +
geom_bar()

library(dplyr)
mpg2 <- mpg %>% arrange(hwy) %>% mutate(id = seq_along(hwy))
ggplot(mpg2, aes(drv, fill = hwy, group = id)) +
geom_bar()
```

(Hint: try adding an outline around each bar with colour = "white")

5. Install the babynames package. It contains data about the popularity of
babynames in the US. Run the following code and fix the resulting graph.
Why does this graph make me unhappy?

- A name can belong to Male or Female
- The scale is bad, because the line of Male is almost parralel to the x-axis 

```{r}
library(babynames)
hadley <- dplyr::filter(babynames, name == "Hadley")
ggplot(hadley, aes(year, n, color = sex)) +
geom_line() +
  theme_bw()
```
Using log 2 scale on the y-axis to see the change
```{r}
library(babynames)
hadley <- dplyr::filter(babynames, name == "Hadley")
ggplot(hadley, aes(year, n, color = sex)) +
geom_line() + 
  scale_y_continuous(trans = 'log2') +
  theme_bw()

```



# 3.6 Surface plot
ggplot2 does not support true 3d surfaces. However, it does support many
common tools for representing 3d surfaces in 2d: contours, coloured tiles and
bubble plots. These all work similarly, differing only in the aesthetic used for
the third dimension.

```{r}
faithfuld %>%
  ggplot(aes(eruptions, waiting, z = density, color= ..level..)) + 
  geom_contour() |
faithfuld %>%
  ggplot(aes(eruptions, waiting, fill = density)) + 
  geom_raster()
```
# 3.7 Drawing Maps
## 3.7.1 Vector Boundaries

Vector boundaries are defined by a data frame with one row for each “corner” of a geographical region like a country, state, or county. It requires four variables:
• `lat` and `long`, giving the location of a point.
• `group`, a unique identifier for each contiguous region.
• `id`, the name of the region.

Separate group and id variables are necessary because sometimes a geographical unit isn’t a contiguous polygon. For example, Hawaii is composed of multiple islands that can’t be drawn using a single polygon.
The following code extracts that data from the built in maps package
using `ggplot2::map data()`. The maps package isn’t particularly accurate or
up-to-date, but it’s built into R so it’s a reasonable place to start.

```{r}
michigan <- map_data('county', 'michigan')
head(michigan)
```

Visualize with `geom_polygon`:
```{r}
michigan %>%
  ggplot(aes(long, lat, group = group)) + 
  geom_polygon(color = 'red') +
  coord_quickmap() 
```
You can also customize the polygon 
```{r}
michigan %>%
  ggplot(aes(long, lat, group = group)) + 
  geom_polygon(fill = NA, color = 'grey50') + 
  coord_quickmap() + 
  theme_void()
```
Note the use of `coord_quickmap()`: it’s a quick and dirty adjustment that
ensures that the aspect ratio of the plot is set correctly


```{r}
#World map
map_data('world') %>%
  ggplot(aes(long, lat, group = group)) + 
  geom_polygon(color = 'white') +
  coord_quickmap()
```
```{r}
#US states
map_data('state') %>%
  ggplot(aes(long, lat, group = group)) + 
  geom_polygon(color = 'white') + 
  coord_quickmap()
```

## 3.7.2 Point metadata
Point metadata connects locations (defined by lat and lon) with other variables. For example, the code below extracts the biggest cities in MI (as of
2006):
```{r}
library(USAboundaries)
mi_cities <- maps::us.cities %>%
filter(country.etc =="MI") %>%
select(-country.etc, lon = long) %>%
arrange(desc(pop))

head(mi_cities)

```
```{r}
ggplot(map_data('county', 'michigan')) + 
  #Draw map
  geom_polygon(aes(long, lat, group = group), fill = NA, color = 'grey50') +
  #For each city in Michigan Country, draw a point with its size represent the population of that city
  #The point is placed at the longtitude `lon` and latitude `lat` of the each city 
  geom_point(data = mi_cities,
             mapping = aes(lon, lat, size = pop), color = 'red') + 
  coord_quickmap()
```

## 3.7.3 Raster Images
Read the book

## 3.7.4 Area metadata

Plot US murder rate 2010

```{r}
library(dslabs)
head(murders)
```

Mapping data

```{r}
state <- map_data('state')
head(state)
```

```{r}
#Convert column murders$state to lower case for joining
murders <- murders %>% mutate(state = tolower(state)) %>% select(state, total)

state <- left_join(state, murders, by = c('region' = 'state'))
state
```

```{r}
state %>%
  ggplot(aes(long, lat, group = group, fill = total)) + 
  geom_polygon(color = 'grey50') +
  scale_fill_distiller(palette = 'Reds') + 
  ggtitle('US murder in 2010') + 
  coord_quickmap() +
  theme_void()
```
# 3.8 Revealing Uncertainty

If you have information about the uncertainty present in your data, whether
it be from a model or from distributional assumptions, it’s a good idea to
display it.

There are four basic families of geoms that can be used for this job,
depending on whether the x values are discrete or continuous, and whether
or not you want to display the middle of the interval, or just the extent:

```{r}
#Data
df <- data.frame(x = 1:3, y = c(18, 11, 16), se = c(1.2, 0.5, 1.0))
```

• Discrete x, range: `geom_errorbar()`, `geom_linerange()`
```{r}
df %>% ggplot(aes(x = x, ymin = y - se, ymax = y + se)) + 
  geom_errorbar() +
  ggtitle('Error bar') |
  
df %>% ggplot(aes(x = x, ymin = y - se, ymax = y + se)) + 
  geom_linerange() +
  ggtitle('Linerange')
```

• Discrete x, range & center: `geom_crossbar()`, `geom_pointrange()`
```{r}
df %>% ggplot(aes(x = x, y = y, ymin = y - se, ymax = y + se)) + 
  geom_crossbar() |
  
df %>% ggplot(aes(x = x, y = y, ymin = y - se, ymax = y + se)) +  
  geom_pointrange()
```

• Continuous x, range: `geom_ribbon()`
```{r}
df %>%
  ggplot(aes(x, y, ymin = y - se, ymax = y + se)) +
  geom_ribbon(fill = 'grey80') +
  geom_line(color = 'red')
```

• Continuous x, range & center: `geom_smooth(stat = "identity")`

```{r}
df %>%
  ggplot(aes(x, y, ymin = y - se, ymax = y + se)) + 
  geom_smooth(stat = 'identity')
```

# 3.9 Weighted Data

Read the book

```{r}
#Linear regression, unweighted (each observation has the same weight)
#min ||x1 - y1||^2 + ||x2 - y2||^2 + ....
tips %>%
  ggplot(aes(total_bill, tip)) +
  geom_point(aes(color = smoker)) +
  geom_smooth(method = 'lm') +
  theme_bw() |

#Linear regression, weighted
#min w1 * ||x1 - y1||^2  + ||x2 - y2||^2 + ....
tips %>%
  #put weight 10 for smoker and weight 1 for non-smoker
  ggplot(aes(total_bill, tip, weight = ifelse(smoker == 'Yes', 10, 1))) +
  geom_point(aes(color = smoker)) +
  geom_smooth(method = 'lm') +
  theme_bw()
```


# 3.10 Diamonds Dataset


```{r}
head(diamonds)
```
## 3.11 Distribution

It is important to experiment with binning to find a revealing view. Never rely on the default parameters to get a revealing view of the distribution.

```{r}
#Diamonds' depth distribution

#Default
diamonds %>%
  ggplot(aes(depth)) +
  geom_histogram() +
  ggtitle('Default: bins = 30') |
  
#Choosing sensible parameters reveal more detail: 
#binwidth = .1
#xlim = c(50, 75)
diamonds %>%
  ggplot(aes(depth)) + 
  geom_histogram(binwidth = .1) + 
  xlim(55, 70) +
  ggtitle('binwidth = 0.1, xlim from 55 to 70')
```
When publishing figures, don’t forget to include information about important parameters (like bin width) in the caption.


If you want to compare the distribution between groups, you have a few
options:

• Show small multiples of the histogram, facet wrap(˜ var).
```{r}
diamonds %>%
  ggplot(aes(depth)) +
  geom_histogram(binwidth = .1) +
  facet_wrap(~ cut) +
  theme_bw()
```

• Use colour and a frequency polygon, geom freqpoly() .

```{r}
diamonds %>%
  ggplot(aes(depth, color = cut)) + 
  geom_freqpoly(size = 1, binwidth = .1) + 
  scale_color_brewer(palette = 'Set1') + 
  xlim(55, 70) +                             #Zoooming
  theme_bw()
```

• Use a “conditional density plot”, geom histogram(position = "fill").

```{r}
diamonds %>%
  ggplot(aes(depth, fill = cut)) + 
  geom_histogram(binwidth = .1, position = 'fill') +
  xlim(55, 70)
```
Both the histogram and frequency polygon geom use the same underlying
statistical transformation: `stat = "bin"`. This statistic produces two output
variables: `count` and `density`. By default, count is mapped to y-position, because it’s most interpretable. The density is the count divided by the total count multiplied by the bin width, and is useful when you want to compare
the shape of the distributions, not the overall size.

```{r}
#Compare the shape of distributions
diamonds %>%
  ggplot(aes(x = depth, y = ..density.., color = cut)) +
  geom_freqpoly(binwidth = .1, size = 1) +
  xlim(50, 75)
```

An alternative to a bin-based visualisation is a density estimate `geom_density()`: Kernel Density Estimation
places a little normal distribution at each data point and
sums up all the curves. It has desirable theoretical properties, but is more
difficult to relate back to the data. Use a density plot when you know that
the underlying density is smooth, continuous and unbounded. You can use
the `adjust` parameter to make the density more or less smooth.

```{r}
diamonds %>%
  ggplot(aes(depth, fill = cut, color = cut)) +
  geom_density(alpha = .2, size = 1) +
  xlim(50, 75) +
  theme_bw()
```
Note that the area of each density estimate is standardised to one so that
you lose information about the relative size of each group.


Compare many distributions:

* boxplot: 5 numbers: min, Q1, median, Q3, max and show outliers 
```{r}
diamonds %>%
  ggplot(aes(cut, depth)) + 
  geom_boxplot()
```
* Violin plot
```{r}
diamonds %>%
  ggplot(aes(cut, depth)) +
  geom_violin()
```
* Dot plot: small dataset

```{r}
mpg %>%
  ggplot(aes(x = class, y = displ)) + 
  geom_dotplot(, binaxis = 'y', stackdir = 'center', dotsize = .5)
```
## 3.11.1 Exercises
1. What binwidth tells you the most interesting story about the distribution
of `carat`?

`binwidth = .1`: There are 2 peeks  
```{r}
p1 <- diamonds %>%
  ggplot(aes(carat)) + 
  geom_histogram(binwidth = .05, color = 'grey80') +
  ggtitle('binwidth = .05')

p2 <- diamonds %>%
  ggplot(aes(carat)) + 
  geom_histogram(binwidth = .075,color = 'grey80') +
  ggtitle('binwidth = .075')

p3 <- diamonds %>%
  ggplot(aes(carat)) + 
  geom_histogram(binwidth = .1, color = 'grey80') +
  ggtitle('binwidth = .1')

p4 <- diamonds %>%
  ggplot(aes(carat)) + 
  geom_histogram(binwidth = .15, color = 'grey80') +
  ggtitle('binwidth = .15')

(p1 | p2) / (p3 | p4)
```

2. Draw a histogram of `price`. What interesting patterns do you see?

- In general, as the price increases, the number of diamonds will decrease exponentially
- There are 3 peeks in the diamonds' price distribution, around 7000, 2000, 4500
- There is a large number of diamonds having price less than 2000$
- Most Diamonds' price is less 10000$, the rest are rare 
```{r}
ggplot(diamonds, aes(price)) +
  geom_histogram(binwidth = 150, color = 'grey80') +
  geom_vline(xintercept = c(700, 2000, 4500), color = 'red') + 
  scale_y_continuous(expand = expansion(mul = c(0, .1))) + 
  theme_light()
```


3. How does the distribution of price vary with clarity?

```{r}
diamonds %>%
  ggplot(aes(price, fill = clarity, color = clarity)) +
  geom_density(alpha = .2, size = .5) + 
  scale_color_brewer(palette = 'Set1') + 
  theme_light() 
  
diamonds %>%
  ggplot(aes(price, fill = clarity)) + 
  geom_histogram(position = 'fill', binwidth = 150) +
  scale_fill_brewer(palette = 'Set2')

diamonds %>%
  ggplot(aes(price, color = clarity)) + 
  geom_freqpoly(binwidth = 150, size = 0.5) +
  scale_color_brewer(palette = 'Dark2') +
  theme_light()


```


4. Overlay a frequency polygon and density plot of depth. What computed
variable do you need to map to y to make the two plots comparable? (You
can either modify `geom_freqpoly()` or `geom_density()`.)

add `y = ..density..` for `geom_freqpoly`
```{r}
diamonds %>%
  ggplot(aes(depth)) +
  geom_freqpoly(aes(y = ..density..), binwidth = .1, color = 'blue') + 
  geom_density(color = 'red', linetype = 'dashed', size = 1) +
  theme_minimal()
```
# 3.12 Dealing with overlapping points

```{r}
#Data 
df <- data.frame(x = rnorm(2000), y = rnorm(2000))
norm <- ggplot(df, aes(x, y)) + xlab(NULL) + ylab(NULL)
```


For small dataset: 
* Chaging Aesthetic: size, shape,...
```{r}
norm + geom_point()

norm + geom_point(shape = 21)  #Hollow circles

norm + geom_point(shape = '.')
```
For larger dataset
* Alpha blending
```{r}
diamonds %>%
  ggplot(aes(carat, price)) + 
  geom_point(alpha = .1)
```
* Jittering + alpha blending: If there is some discrete in the data

```{r}
diamonds %>%
  ggplot(aes(cut, price)) +
  geom_jitter(width = .2, alpha = .1)
```
* 2D density
```{r}
#Hexagon 
diamonds %>%
  ggplot(aes(carat, price)) + 
  geom_hex() +
  theme_light() 

#2D kernel density estimation
diamonds %>%
  ggplot(aes(carat, price)) + 
  geom_density_2d_filled()
#2D bin
diamonds %>%
  ggplot(aes(carat, price)) + 
  geom_bin2d() +
  theme_light()
```

# 3.13 Statistical Summary
Let’s start with a couple of examples with the diamonds data. The first
example in each pair shows how we can count the number of diamonds in
each bin; the second shows how we can compute the average price.

```{r}
#The number of diamonds for each color
diamonds %>%
  ggplot(aes(color)) + 
  geom_bar(fill = 'tomato') +
  theme_light()
```
```{r}
#The average price for each color

#Using stat_summary
ggplot(diamonds, aes(color, price)) +
  stat_summary(fun = 'mean', geom = 'col', 
               fill = 'tomato') +
  theme_light()

#Equivalently, Using geom_bar with summary_bin
ggplot(diamonds, aes(color, price)) +
  geom_bar(stat = 'summary_bin', fun = mean, 
           fill = 'tomato') +
  theme_light()
```
2D summary
```{r}
#Median price for different depth and table
diamonds %>%
  ggplot(aes(table, depth)) +
  stat_summary_2d(aes(z = price), geom = 'hex', fun = median) +
  theme_light()

#Equivalent
diamonds %>%
  ggplot(aes(table, depth)) +
  geom_hex(aes(z = price), stat = 'summary_2d', fun = median) + 
  theme_light()
```



# 3.14 Add-on Packages

Read book packages that extend ggplot2: ggforce, ggalt, ...

