---
title: "The basic of Data Exploration and Visualization with R"
author: "Pikachu"
date: "11/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Exercise 1: Installing and loading tidyverse
```{r}
# install.packages('tidyverse')
library(tidyverse)
```

# Exercise 2: Loading and examining a dataset
Read the contents of a file called
`Crime_Data.csv`. This file contains approximately 481,000 crime reports from
Seattle, WA covering a span of approximately 10 years.

```{r}
crime <- read_csv('./IntroR/Data/Crime_data.csv', 
                  col_types = cols(`Report Number` = col_double(),
                                   `Occurred Date` = col_date('%m/%d/%Y'),
                                   `Occurred Time` = col_double(),
                                   `Reported Date` = col_date('%m/%d/%Y'),
                                   `Reported Time` = col_double(),
                                   `Crime Subcategory` = col_character(),
                                   `Primary Offense Description` = col_character(),
                                   Precinct = col_character(),
                                   Sector = col_character(),
                                   Beat = col_character(),
                                   Neighborhood = col_character()),
                  na = c('UNKNOWN')
                  )
crime

# View the tibble in a new tab
# View(crime)
```
Limit the columns to the following: 
Reported Date, Crime Subcategory , Primary Offense Description, Precinct, Sector, Beat, and Neighborhood. 

```{r}
variables <- c('Reported Date', 'Crime Subcategory', 'Primary Offense Description',
               'Precinct', 'Sector', 'Beat', 'Neighborhood')
dfCrime <- crime %>% select(all_of(variables))

dfCrime


```

# Exercise 3: Filtering a dataset

# Exercise 4: Grouping and summarizing a dataset

# Exercise 5: Plotting a dataset

# Exercise 6: Graphing burglaries by month and year

