---
title: "Names and Values"
author: "Pikachu"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(lobstr)
```

# 2.1 Introduction

whenever you modify a vector, you’re almost certainly creating a new, modified vector

# 2.2 Binding basics

>Access object's idetifier with `**lobstr::obj_addr**`

```{r}
# x and y poin to the same address
x <- c(1, 2, 3)

y <- x

x %>% obj_addr()
y %>% obj_addr()
```

## 2.2.1 Non-syntactic names

```{r}
`this is a non syntatic name` <- 3
`this is a non syntatic name`
```

## 2.2.3 Exercises

>1. Explain the relationship between `a`, `b`, `c` and `d` in the following code:

```{r include = F}
a <- 1:10
b <- a
c <- b
d <- 1:10
```
`a`, `b` and `c` point to the same memory address

```{r}
obj_addr(a)
obj_addr(b)
obj_addr(c)
obj_addr(d)
```
```{r}
obj_addrs(list(a, b, c, d)) 

```



>2. The following code accesses the mean function in multiple ways. Do they all point to the same underlying function object? Verify this with `lobstr::obj_addr()``.

```{r include = F}
mean
base::mean
get("mean")
evalq(mean)
match.fun("mean")
```

They all point to the same underlying function object(the same memory address)

```{r}
obj_addr(mean)
obj_addr(base::mean)
obj_addr(get('mean')) 
obj_addr(evalq(mean))
obj_addr(match.fun('mean'))

```



>3. By default, base R data import functions, like `read.csv()``, will automatically convert non-syntactic names to syntactic ones. Why might this be problematic? What option allows you to suppress this behaviour?

Use `tibble`

>4. What rules does `make.names()` use to convert non-syntactic names into syntactic ones?

Read `?make.names()`

>5. I slightly simplified the rules that govern syntactic names. Why is `.123e1`` not a syntactic name? Read `?make.names` for the full details.

# 2.3 Copy-on-modify

R objects are immutable. Consider this example:

```{r}
x <- c(1, 2, 3)
y <- x

# x and y share the same memory address
obj_addrs(list(x, y)) %>% unique()

# modify y
y[1] <- -1

# y changed
y
# x does not change
x

# now the address of y changed
obj_addrs(list(x, y))
```

```{r}
a <- c(1, 2, 3)
obj_size(a)
b <- a 
# a and b reference to the same object
# so obj_size(a, b) is equal to obj_size(a)
obj_size(a, b)

b[[1]] <- 999

# when you modify, copy-on-modify will occur, create a new vector
# Now a and b do not share reference
# obj_size(a, b) will be doubled
obj_size(a, b)
```


## 2.3.1 `tracemem()`

From then on, whenever that object is copied, `tracemem()` will print a message telling you which object was copied, its new address, and the sequence of calls that led to the copy:

```{r}
x <- c(1, 2, 3)

tracemem(x)

y <- x

y[1] <- 3


```
>`untracemem()` is the opposite of `tracemem`

## 2.3.2 Function calls

## 2.3.3 List

## 2.3.4 Data frames

## 2.3.5 Character vectors

## 2.3.6 Exercises

>1. Why is `tracemem(1:10)` not useful?

It is not useful because when `1:10` is called, an object with an address in memory in created. But is is not bounded to a name. Therefore, this object can't not be called or manipulated from R.

```{r}
obj_addr(1:10)
```


>2. Explain why `tracemem()` shows two copies when you run this code. Hint: carefully look at the difference between this code and the code shown earlier in the section.

```{r}
x <- c(1L, 2L, 3L)
tracemem(x)

x[[3]] <- 4
```
Initially the vector `x` has integer type. The replacement call assigns a double to the third element of `x`, which triggers copy-on-modify.

>3. Sketch out the relationship between the following objects:

```{r }
a <- 1:10
b <- list(a, a)
c <- list(b, a, 1:10)

# reference tree
ref(c)
```

>4. What happen when you run this code?

```{r include=F}
x <- list(1:10)
x[[2]] <- x
```

# 2.4 Object size 

You can find out how much memory an object takes with `lobstr::obj_size()`:

```{r}
obj_size(letters)

obj_size(ggplot2::diamonds)
```
Since the elements of lists are references to values, the size of a list might be much smaller than you expect:

```{r}
x <- rnorm(1e6)

obj_size(x)

y <- list(x, x, x)
obj_size(y)
```
`y` is only `r obj_size(y) - obj_size(x)` bigger than `x`. That's the size of an empty list with 3 elements:

```{r}
list(NULL, NULL, NULL) %>% obj_size()
```

Similarly, because R uses a global string pool character vectors take up less memory than you might expect: repeating a string 100 times does not make it take up 100 times as much memory.

```{r}
fruit <- 'banana'

obj_size(fruit)

fruit %>% rep(100) %>% obj_size()
```
References also make it challenging to think about the size of individual objects. `obj_size(x) + obj_size(y)` will only equal `obj_size(x, y)` if there are no shared values.

```{r}
x <- rnorm(100)

y <- x

# object size of x is 848B
obj_size(x)

# object size of x and y is also 848B, because y references x
obj_size(x, y)
```
nally, R 3.5.0 and later versions have a feature that might lead to surprises: ALTREP, short for __alternative representation__. This allows R to represent certain types of vectors very compactly. The place you are most likely to see this is with : because instead of storing every single number in the sequence, R just stores the first and last number. This means that every sequence, no matter how large, is the same size:

```{r}
obj_size(1:10)

obj_size(1:1e9)
```

## 2.4.1 Exercises
Read the solution book. This is very challenging.

# 2.5 Modify-in-place
## 2.5.1 Objects with single binding
## 2.5.2 Environment

Environment are always modified-in-place
```{r}
e1 <- rlang::env(a = 1, b = 2, c = 3)

e2 <- e1

e2$b <- -1

# e1$b will also be changed to -1
e1$b
```


# 2.6 Unbindding and the garbage collector

