---
title: "Introduction"
author: "Pikachu"
date: "11/20/2020"
output: html_document
---
Reference: [tibble](https://tibble.tidyverse.org/articles/tibble.html)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```

# Creating
__tibble()__ is a nice way to create data frames. It encapsulates best practices for data frames:
- It never changes an input’s type (i.e., no more `stringsAsFactors = FALSE!`).

```{r}
tibble(values = letters)
```

This makes it easier to use with list-columns:

```{r}
tibble(x = 1:3, y = list(1:5, 1:10, 1:20))
```
- It never adjusts the names of variables:

```{r}
data.frame(`Sepal length` = 1)
```

```{r}
tibble(`Sepal Length` = 1)
```

- It evaluates its arguments lazily and sequentially:

```{r}
tibble(x = 1:5, y = x * 3)
```

- It never uses row.names(). The whole point of tidy data is to store variables in a consistent way. So it never stores a variable as special attribute.

- It only recycles vectors of length 1. This is because recycling vectors of greater lengths is a frequent source of bugs.

# Coercion
Using **`as_tibble`**

# Tibbles vs Data frames

There are three key differences between tibbles and data frames: printing, subsetting, and recycling rules.

## Printing
When you print a tibble, it only shows the first ten rows and all the columns that fit on one screen. It also prints an abbreviated description of the column type, and uses font styles and color for highlighting:

```{r}
as_tibble(iris)
```


You can control the default appearance with options:

- `options(tibble.print_max = n, tibble.print_min = m)`: if there are more than n rows, print only the first m rows. Use `options(tibble.print_max = Inf)` to always show all rows.

- `options(tibble.width = Inf)` will always print all columns, regardless of the width of the screen.

## Subsetting

Tibbles are quite strict about subsetting. `[` always returns another tibble.
Contrast this with a data frame: sometimes [ returns a data frame and sometimes it just returns a vector:

To extract a single column use `[[` or `$`:

```{r}
iris <- as_tibble(iris)

iris['Species']          # This returns a tibble

iris$Species             # This returns a vector
```

## Recyling

When constructing a tibble, only values of length 1 are recycled. The first column with length different to one determines the number of rows in the tibble, conflicts lead to an error. This also extends to tibbles with zero rows, which is sometimes important for programming:

```{r}
tibble(x = 1, y = 1:3)
```

```{r}
tibble(x = 1:2, y = 1:3)
```

